Keep Your Laziness In Check

### 2017-11-20 slides.tex
More edits to slide content

\note[itemize]{
  \item As a micro example to motivate why thinking about laziness can be
        non-trivial, let's look at this queue
  \item In this slide, we use call-by-name ML for simplicity of presentation,
        but it can be implemented in OCaml using the Lazy module
  \item This has amortized O(1) performance cost.
  \item This queue is lazy as long as you don't empty the front, then it is
        fully lazy in the structure of the back list. It only forces the spine
        of the back list when the front list is emptied.
}

@@

\note[itemize]{
\item PBRT is a technique for randomized testing which allows users to write
  down an executable specification for a function and test whether it holds for
  randomly generated inputs.
\item QuickCheck gives functional correctness
\item But laziness can't be observed by QuickCheck
\item Say that laziness is not generally considered
      part of functional correctness much like how asymptotic runtime
      is not part of functional correctness because they are not directly
      observable
\item What if we make it observable? Then it's just part of functional correctness!
}

@@

\begin{frame}[fragile]
\frametitle{\texttt{ListDemand}}
\begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil
------------------------------------
type 'a thunk = T | E of 'a

@@

\begin{frame}[fragile]
\frametitle{\texttt{TreeDemand}}
\begin{minted}{ocaml}
type 'a lazyTree =
  | Node of ('a lazyTree Lazy.t *
             'a Lazy.t *
             'a lazyTree Lazy.t)
  | Leaf

@@

\begin{frame}[fragile]
\frametitle{\texttt{ListDemand} vs \texttt{TreeDemand}}
\begin{minted}{ocaml}
type 'd listDemand =
  | ConsD of ('d thunk *
              'd listDemand thunk)
  | NilD

@@

@@

\begin{frame}[fragile]
\frametitle{Have questions?}
Come ask us about \dots
\begin{itemize}
\item Higher-order specifications
\item Generic programming for any arity and datatype
\item Random generation of partial demand contexts
\item Efficiency of our implementation
\item Details of instrumentation
\item What language are you actually implementing this in? \\ (Hint: it's not ML)
\item[] \dots
\end{itemize}
\end{frame}

>>>>

\note[itemize]{
  \item As a micro example to motivate why thinking about laziness can be
        non-trivial, let's look at this queue
  \item This has amortized O(1) performance cost.
  \item This queue is lazy as long as you don't empty the front, then it is
        fully lazy in the structure of the back list. It only forces the spine
        of the back list when the front list is emptied.
}

@@

\note[itemize]{
\item PBRT is a technique for randomized testing which allows users to write
  down an executable specification for a function and test whether it holds for
  randomly generated inputs.
\item QuickCheck gives functional correctness
\item But laziness can't currently be observed by tools like QuickCheck
\item Laziness is not generally considered part of functional correctness much
  like how asymptotic runtime is not part of functional correctness because they
  are not directly observable
\item What if we make it observable? Then it's just part of functional correctness!
}

@@

\begin{frame}[fragile]
  \frametitle{Reifying demand on lazy lists}
\begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil
------------------------------------
type 'a thunk = T | E of 'a

@@

\begin{frame}[fragile]
  \frametitle{Reifying demand on lazy trees}
\begin{minted}{ocaml}
type 'a lazyTree =
  | Node of ('a lazyTree Lazy.t *
             'a Lazy.t *
             'a lazyTree Lazy.t)
  | Leaf

@@

\begin{frame}[fragile]
  \frametitle{Comparing demands on lazy lists and trees}
\begin{minted}{ocaml}
type 'd listDemand =
  | ConsD of ('d thunk *
              'd listDemand thunk)
  | NilD

@@

\begin{frame}[fragile]
  \frametitle{Recalling \texttt{deQ}}
  \begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil

@@

\begin{frame}[fragile]
\frametitle{Have questions?}
Come ask us about \dots
\begin{itemize}
\item Higher-order specifications
\item Generic programming for any arity and datatype
\item Random generation of partial demand contexts
\item Efficiency of our implementation
\item Details of instrumentation
\item What language are you actually implementing this in? \\ (Hint: it's not ML)
\item[] \ \ $\vdots$
\end{itemize}
\end{frame}
